Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{ I thank Lara Andermann, Leonie Bielefeld, Daniel Bursian, Davide Cantoni, Emilio Esguerra, Filip Milojevic, Johannes Vorbach and Matthias Weigand for helpful comments and feedback. }

### 2023-06-27 paper/paper.tex
#29 finish robustness (ext validity)

Having discussed issues with internal validity now, I will briefly talk about external validity. The hypothesis I test is a prediction of the theory of conflict-driven evolution. While the theory itself can be applied widely --- \cite{levine2021} use it to explain stylised facts about pre-modern Europe, pre-modern India, the present-day United Kingdom and Spain, and 20th-century China --- my empirical results only test whether the theory can explain what happens in my dataset.

@@

>>>>

Having discussed issues with internal validity now, I will briefly talk about external validity. We cannot infer from my results that the theory of conflict-driven evolution is true or false. I test a general prediction of the theory in the Holy Roman Empire between 1300 and 1800, which as a setting is fairly unique. It features a vast array of fragmented states that compete and go to war with each other, but are also part of the same empire and likely to share many institutions. Religious conflict peaks in the sample period, with the Thirty Years War taking place from 1618-1648, and alliances are often based on the shared confession of rulers. Spain, by contrast, was a large monolithic state by the end of the Reconquista in 1492, in which the Protestant reformation never really took off.

Based on these unique features of the setting we should be wary even of generalising my results to the rest of Europe.\footnote
{
\cite{schoenholzer2022} fill this gap partially, given that their sample comprises all of Europe and they at least find a similar overall effect of switching states on long-run development.
}
Nonetheless, the results have value outside of understanding the history of the Holy Roman Empire: as a puzzle piece. A main objective of this paper was to test a theory. Theories become more widely accepted when researchers repeatedly test and confirm their predictions; they are discredited when their predictions stubbornly disagree with reality. Even the result that there is no evidence for the prediction is valuable because it points towards limitations of the theory.

@@

What do we learn from this paper?
